### 2024-06-18 analysis/models-for-symposium-workshop.qmd
Add models from symposium workshop

---
title: "Symposium models"
---

```{r}
#| label: setup
#| include: false

if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE, warning = FALSE, message = FALSE
)

knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)
```

```{r}
#| label: libraries-data

library(tidyverse)
library(haven)
library(modelsummary)
library(tinytable)
library(marginaleffects)
library(MASS)
library(conflicted)

conflicts_prefer(dplyr::select)

theme_set(theme_minimal())

config_modelsummary(factory_default = 'tinytable')

derog <- read_stata(here::here("data", "raw_data", "JHR Symposium HR and DB 4 30 24 stata data.dta")) |>
  zap_formats() |>
  zap_label() |>
  mutate(
    c5_public_transport = factor(c5_public_transport,
      levels = 0:2,
      labels = c("No measures", "Recommend closing", "Require closing"),
      ordered = TRUE
    ),
    c6_stay_at_home = factor(c6_stay_at_home,
      levels = 0:3,
      labels = c(
        "No measures", 
        "Recommend not leaving house", 
        "Require not leaving house, with exceptions", 
        "Require not leaving house, with minimal exceptions"
      ),
      ordered = TRUE
    ),
    c7_internal_movement = factor(c7_internal_movement,
      levels = 0:2,
      labels = c("No measures", "Recommend to not travel", "Restrictions in place"),
      ordered = TRUE
    ),
    e1_income_support = factor(e1_income_support,
      levels = 0:2,
      labels = c(
        "No income support", 
        "Government replaces less than 50% of lost salary", 
        "Government replaces more than 50% of lost salary"
      ),
      ordered = TRUE
    ),
    e2_debt_relief = factor(e2_debt_relief,
      levels = 0:2,
      labels = c("No debt relief", "Narrow relief", "Broad relief"),
      ordered = TRUE
    ),
    pandem_discrim = factor(pandem_discrim,
      levels = c("None", "Minor", "Moderate", "Major"),
      ordered = TRUE
    ),
    pandem_ndrights = factor(pandem_ndrights,
      levels = c("None", "Major"),
      ordered = TRUE
    ),
    pandem_abusive = factor(pandem_abusive,
      levels = c("None", "Minor", "Moderate", "Major"),
      ordered = TRUE
    ),
    pandem_nolimit = factor(pandem_nolimit,
      levels = c("None", "Moderate", "Major"),
      ordered = TRUE
    ),
    pandem_media = factor(pandem_media,
      levels = c("None", "Minor", "Moderate", "Major"),
      ordered = TRUE
    )
  ) |>
  mutate(
    panbackdichot = panbackdichot - 1,
    panbackdichot_bin = as.logical(panbackdichot)
  ) |> 
  mutate(
    across(
      c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
      list(z = ~ as.numeric(scale(.)))
    )
  )

gof_map <- tribble(
  ~raw,          ~clean,                 ~fmt, ~omit,
  "nobs",        "N",                    0,    FALSE
)

coef_map <- c(
  "(Intercept)" = "Intercept",
  "derogation_ineffect" = "Derogation in effect",
  "panbackdichot" = "Pandemic backsliding (PanBack), dichotomous",
  "derogation_ineffect:panbackdichot" = "Derogation in effect × Pandemic backsliding",
  "panback" = "Pandemic backsliding (PanBack)",
  "pandem" = "Pandemic violations of democratic standards (PanDem)",
  "new_cases" = "New cases",
  "new_cases_z" = "New cases (standardized)",
  "new_deaths" = "New deaths",
  "new_deaths_z" = "New deaths (standardized)",
  "cumulative_cases" = "Cumulative cases",
  "cumulative_cases_z" = "Cumulative cases (standardized)",
  "cumulative_deaths" = "Cumulative deaths",
  "cumulative_deaths_z" = "Cumulative deaths (standardized)",
  "v2x_rule" = "Rule of law index",
  "No measures|Recommend to not travel" = "No measures | Recommend to not travel",
  "Recommend to not travel|Restrictions in place" = "Recommend to not travel | Restrictions in place",
  "No measures|Recommend closing" = "No measures | Recommend closing",
  "Recommend closing|Require closing" = "Recommend closing | Require closing",
  "No measures|Recommend not leaving house" = "No measures | Recommend not leaving house",
  "Recommend not leaving house|Require not leaving house, with exceptions" = "Recommend not leaving house| Require not leaving house, with exceptions",
  "Require not leaving house, with exceptions|Require not leaving house, with minimal exceptions" = "Require not leaving house, with exceptions | Require not leaving house, with minimal exceptions",
  "No income support|Government replaces less than 50% of lost salary" = "No income support | Government replaces less than 50% of lost salary",
  "Government replaces less than 50% of lost salary|Government replaces more than 50% of lost salary" = "Government replaces less than 50% of lost salary | Government replaces more than 50% of lost salary",
  "No debt relief|Narrow relief" = "No debt relief | Narrow relief",
  "Narrow relief|Broad relief" = "Narrow relief | Broad relief",
  "None|Minor" = "None | Minor",
  "Minor|Moderate" = "Minor | Moderate",
  "Moderate|Major" = "Moderate | Major",
  "None|Moderate" = "None | Moderate"
)

tidy_custom.polr <- function(x, ...) {
  s <- lmtest::coeftest(x)
  out <- data.frame(
    term = row.names(s),
    p.value = s[, "Pr(>|t|)"])
  out
}
```

# 1: Explaining COVID-19 Derogations

```{r}
#| label: tbl-derogations
#| include: true
#| tbl-cap: "Explaining COVID-19 Derogations"

model1 <- glm(
  iccpr_derogation_filed ~ panback + new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  family = binomial(link = "logit")
)

model2 <- glm(
  iccpr_derogation_filed ~ pandem + new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  family = binomial(link = "logit")
)

# Save models for II paper appendix
# saveRDS(
#   lst(models = list("Model 1" = model1, "Model 2" = model2), gof_map, coef_map), 
#   "~/Desktop/determinants_models.rds"
# )

modelsummary(
  list("Model 1" = model1, "Model 2" = model2), 
  exponentiate = TRUE,
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
  coef_map = coef_map,
  gof_map = gof_map,
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Logistic regression models; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  )
)
```

```{r}
#| label: fig-pred-derogations
#| include: true

plot_predictions(model1, condition = "panback") +
  labs(y = "Predicted probability of derogation")

plot_predictions(model2, condition = "pandem") +
  labs(y = "Predicted probability of derogation")
```

# 2: Explaining COVID-19 Restrictions

```{r}
#| label: tbl-restrictions
#| include: true
#| tbl-cap: "Explaining COVID-19 Restrictions"

model3 <- polr(
  c7_internal_movement ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model4 <- polr(
  c5_public_transport ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model5 <- polr(
  c6_stay_at_home ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

modelsummary(
  list(
    "Model 3<br>(Internal movement)" = model3, 
    "Model 4<br>(Public transportation)" = model4, 
    "Model 5<br>(Stay at home)" = model5
  ), 
  exponentiate = TRUE,
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
  coef_map = coef_map,
  gof_map = gof_map,
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Ordered logistic regression models; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  )
)
```

```{r}
#| label: fig-pred-restrictions
#| include: true
#| fig-width: 7

plot_predictions(model3, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$c7_internal_movement)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\ninternal movement measures",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model4, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$c5_public_transport)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\ntransportation measures",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model5, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$c6_stay_at_home)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nstay at home measures",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

# 3: Explaining COVID-19 Economic Policies

```{r}
#| label: tbl-economic-policies
#| include: true
#| tbl-cap: "Explaining COVID-19 Economic Policies"

model6 <- polr(
  e1_income_support ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model7 <- polr(
  e2_debt_relief ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

modelsummary(
  list(
    "Model 6<br>(Income support)" = model6, 
    "Model 7<br>(Debt relief)" = model7
  ), 
  exponentiate = TRUE,
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
  coef_map = coef_map,
  gof_map = gof_map,
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Ordered logistic regression models; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  )
)
```

```{r}
#| label: fig-pred-economic-policies
#| include: true
#| fig-width: 6
#| fig-height: 5

plot_predictions(model6, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$e1_income_support))), ncol = 1) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nincome support",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model7, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$e2_debt_relief))), ncol = 1) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\ndebt relief",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

# 4: Explaining COVID-19 Human Rights Violations

```{r}
#| label: tbl-human-rights
#| include: true
#| tbl-cap: "Explaining COVID-19 Human Rights Violations"

model_discrim <- polr(
  pandem_discrim ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model_ndrights <- glm(
  pandem_ndrights ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  family = binomial(link = "logit")
)

model_abusive <- polr(
  pandem_abusive ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model_nolimit <- polr(
  pandem_nolimit ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

model_media <- polr(
  pandem_media ~ derogation_ineffect*panbackdichot + 
    new_cases_z + new_deaths_z + cumulative_cases_z + cumulative_deaths_z + v2x_rule,
  data = derog,
  method = "logistic",
  Hess = TRUE
)

modelsummary(
  list(
    "Model 8<br>(Discriminatory policy)" = model_discrim, 
    "Model 9<br>(Non-derogable rights)" = model_ndrights,
    "Model 10<br>(No time limit measures)" = model_abusive,
    "Model 11<br>(Abusive enforcement)" = model_nolimit,
    "Model 12<br>(Media restrictions)" = model_media
  ), 
  exponentiate = TRUE,
  estimate = "{estimate}{stars}",
  statistic = "({std.error})",
  coef_map = coef_map,
  gof_map = gof_map,
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Models 8, 10, 11, and 12 are ordered logistic regression models; Model 9 is a logistic regression model; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  )
)
```

```{r}
#| label: fig-pred-human-rights
#| include: true

plot_predictions(model_discrim, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$pandem_discrim)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\ndiscriminatory policy",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model_ndrights, condition = c("panbackdichot", "derogation_ineffect")) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nviolating non-derogable rights",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model_nolimit, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$pandem_nolimit)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nno time limits",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model_abusive, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$pandem_abusive)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nabusive enforcement",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_predictions(model_media, condition = c("panbackdichot", "derogation_ineffect")) +
  facet_wrap(vars(factor(group, levels = levels(derog$pandem_media)))) +
  labs(
    x = "Derogation in effect",
    y = "Predicted probability of\nmedia restrictions",
    color = "panback"
  ) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

# 5: Explaining Non-Derogation Human Rights Treaty Actions during COVID-19

```{r}
#| label: tbl-treaty-actions
#| include: true
#| tbl-cap: "Explaining Non-Derogation Human Rights Treaty Actions during COVID-19"

model8 <- glm(
  noniccprtreatyactionsdichotomous ~ panback + new_cases + new_deaths + cumulative_cases + cumulative_deaths + v2x_rule,
  data = derog,
  family = binomial(link = "logit")
)

model9 <- glm(
  noniccprtreatyactionsdichotomous ~ pandem + new_cases + new_deaths + cumulative_cases + cumulative_deaths + v2x_rule,
  data = derog,
  family = binomial(link = "logit")
)

modelsummary(
  list("Model 13" = model8, "Model 14" = model9), 
  exponentiate = TRUE,
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
  coef_map = coef_map,
  gof_map = gof_map,
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Logistic regression models; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  )
)
```

```{r}
#| label: fig-pred-treaty-actions
#| include: true

plot_predictions(model8, condition = "panback") +
  labs(y = "Predicted probability of derogation")

plot_predictions(model9, condition = "pandem") +
  labs(y = "Predicted probability of derogation")
```