### 2024-06-24 analysis/appendix.qmd
Run all models with cmdstanr and targets

---
title: Appendix
toc: false
page-layout: full
---

TODO