### 2024-06-24 analysis/misc-analysis.qmd
Add maps

---
title: "Other analysis"
format:
  html:
    code-tools: false
    include-in-header: 
      - text: |
          <style>
          .table th {
            font-weight: bold
          }
          </style>
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  # echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
  fig.width = 6, fig.height = (6 * 0.618), out.width = "80%", 
  fig.align = "center", fig.retina = 3,
  collapse = TRUE
)

options(digits = 3, width = 120, dplyr.summarise.inform = FALSE)
```

```{r}
#| label: libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(scales)
library(sf)
library(tinytable)
library(targets)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

tar_load(map_with_data)
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
```

```{r}
#| label: fig-derogation-map
#| fig-width: 6.5
#| fig-height: 3.5
#| out-width: 100%
#| fig-cap: "ICCPR pandemic-related derogations"

ggplot() +
  geom_sf(
    data = map_with_data, aes(fill = derogations_1plus),
    linewidth = 0.1, color = "white"
  ) +
  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
    high = clrs[1], na.value = "grey85",
    breaks = c(1, 5, 10, 15, 20),
    limits = c(1, 20),
    name = "Derogations filed: ",
    guide = guide_colorsteps(barwidth = 7, barheight = 0.4, title.vjust = 1)
  ) +
  theme_pandem() +
  theme(
    panel.border = element_blank(),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = rel(0.6)),
    legend.justification = "center"
  )
```

```{r}
#| label: fig-panback-map
#| fig-width: 6.5
#| fig-height: 3.5
#| out-width: 100%
#| fig-cap: "Average Pandemic Backsliding Index from V-Dem from March 2020–June 2021"

ggplot() +
  geom_sf(
    data = map_with_data, aes(fill = avg_panback_cut),
    linewidth = 0.05, color = "white"
  ) +
  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
  scale_fill_manual(
    values = clrs[c(8, 7, 2, 1)],
    na.value = "grey85",
    breaks = levels(map_with_data$avg_panback_cut),
    name = "Average Pandemic Backsliding Index"
  ) +
  theme_pandem() +
  theme(
    panel.border = element_blank(),
    legend.position = "bottom",
    legend.justification = "center"
  )
```