### 2024-07-10 analysis/analysis-derogations.qmd
PHEW recreate google doc version + targetsify + renvify + clean up website

---
title: "Explaining derogations"
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  fig.width = 6, fig.height = (6 * 0.618), out.width = "80%",
  fig.align = "center", fig.retina = 3,
  collapse = TRUE
)

options(digits = 3, width = 120, dplyr.summarise.inform = FALSE)
```

```{r}
#| label: libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(tidybayes)
library(modelsummary)
library(scales)
library(glue)
library(ggtext)
library(tinytable)
library(targets)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

tar_load(c(m_derogations, m_tbl_derogations, action_state_type))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(diagnostic_functions), .GlobalEnv))
invisible(list2env(tar_read(helper_functions), .GlobalEnv))
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
```

# Model details

## Formal model specification

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$}} \\
\text{Treaty action}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\pi_{it} =&\ \beta_0 + \beta_1\ \text{PanBack}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Rule of law index}_{it}\ + \beta_7\ \text{Week number}_{it} \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 7} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3)
\end{aligned}
$$

## Priors

```{r}
#| label: figure-prior
#| fig-cap: "Density plot of prior distribution for model parameters"
#| fig-width: 3.5
#| fig-height: 2.5
#| out-width: "50%"

ggplot() +
  stat_function(
    geom = "area",
    fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3),
    fill = clrs[2]
  ) +
  xlim(c(-20, 20)) +
  labs(x = "βs") +
  facet_wrap(vars("β: Student t(ν = 1, µ = 0, σ = 3)")) +
  theme_pandem(prior = TRUE)
```

## Simplified R code

```r
brm(
  bf(outcome ~ panback +
      new_cases_z + cumulative_cases_z +
      new_deaths_z + cumulative_deaths_z +
      v2x_rule + year_week_num),
  family = bernoulli(),
  prior = c(
    prior(student_t(1, 0, 3), class = Intercept),
    prior(student_t(1, 0, 3), class = b)),
  ...
)
```

## Model evaluation

```{r}
params_to_show <- c("b_Intercept", "b_panback", "b_new_cases_z", "b_v2x_rule")
```

::: {.panel-tabset}

### Derogation filed

```{r}
plot_trace(m_derogations$m_derogations_panback, params_to_show)
```

```{r}
plot_trank(m_derogations$m_derogations_panback, params_to_show)
```

```{r}
plot_pp(m_derogations$m_derogations_panback)
```

### Other treaty action

```{r}
plot_trace(m_derogations$m_other_panback, params_to_show)
```

```{r}
plot_trank(m_derogations$m_other_panback, params_to_show)
```

```{r}
plot_pp(m_derogations$m_other_panback)
```

:::

# Results

```{r}
#| label: calc-derog-coefs

coef_lookup <- tribble(
  ~coef, ~coef_nice,
  "b_panback", "Pandemic backsliding index (PanBack)",
  "b_v2x_rule", "Rule of law index",
  "b_new_cases_z", "New cases (standardized)",
  "b_cumulative_cases_z", "Cumulative cases (standardized)",
  "b_new_deaths_z", "New deaths (standardized)",
  "b_cumulative_deaths_z", "Cumulative deaths (standardized)"
) |>
  mutate(coef_nice = fct_inorder(coef_nice))

m_derog_draws <- m_derogations$m_derogations_panback |>
  gather_draws(`^b_.*`, regex = TRUE) |>
  filter(.variable %in% coef_lookup$coef) |>
  left_join(coef_lookup, by = join_by(.variable == coef))

derog_coefs <- m_derog_draws |>
  mutate(.value_exp = exp(.value)) |>
  group_by(.variable, coef_nice) |>
  reframe(
    post_medians = median_hdci(.value_exp, .width = 0.95),
    p_gt_0 = sum(.value_exp > 1) / n()
  ) |>
  unnest(post_medians) |>
  mutate(
    y_nice = fmt_coef(y),
    y_nice_html = fmt_coef(y, html = TRUE)
  ) |>
  mutate(
    p_lt_0 = 1 - p_gt_0,
    p_gt = fmt_p_inline(p_gt_0, "gt"),
    p_lt = fmt_p_inline(p_lt_0, "lt"),
    p_gt_html = fmt_p_inline(p_gt_0, "gt", html = TRUE),
    p_lt_html = fmt_p_inline(p_lt_0, "lt", html = TRUE)
  ) |>
  mutate(
    p_d = if_else(y > 1, p_gt, p_lt),
    p_d_html = if_else(y > 1, p_gt_html, p_lt_html),
    plot_label = glue("{y_nice_html}; {p_d_html}")
  ) |>
  mutate(or_pct = label_percent(accuracy = 1)(abs(1 - y)))
```

## Coefficient plot

```{r}
#| label: figure-derog-coefs
#| fig-width: 6
#| fig-height: 4.5
#| fig-cap: "Odds ratios for coefficients from logistic regression model predicting the probability of derogation from the ICCPR"
#| out-width: 100%

m_derog_draws |>
  mutate(.value = exp(.value)) |>
  ggplot(aes(x = .value, y = fct_rev(coef_nice))) +
  stat_pointinterval() +
  geom_vline(xintercept = 1, linewidth = 0.25, linetype = "21") +
  geom_richtext(
    data = derog_coefs, aes(x = y, label = plot_label),
    size = 2.7, nudge_y = 0.35, label.size = 0.1, label.colour = "grey50"
  ) +
  scale_x_log10() +
  labs(
    x = "Odds ratio", y = NULL,
    caption = str_wrap(
      glue(
        "Point shows posterior median;", "
        thick lines show 80% credible interval;",
        "thin black lines show 95% credible interval",
        .sep = " "
      ),
      width = 60
    )
  ) +
  theme_pandem() +
  theme(panel.grid.major.y = element_blank())
```

## Complete table of results

```{r}
#| label: table-results-full-derogations
#| tbl-cap: "Complete results from models showing predictors of derogations (H~1~)"

notes <- paste(
  "Note: Estimates are median posterior log odds from logistic regression models;",
  "95% credible intervals (highest density posterior interval, or HDPI) in brackets."
)

m_tbl_derogations |>
  set_names("Derogation filed", "Other action") |>
  modelsummary(
    estimate = "{estimate}",
    statistic = "[{conf.low}, {conf.high}]",
    coef_map = coef_map,
    gof_map = gof_map,
    output = "tinytable",
    fmt = fmt_significant(2),
    notes = c(notes),
    width = c(0.4, rep(0.3, 2))
  ) |>
  group_tt(j = list("ICCPR action" = 2:3)) |>
  style_tt(j = 2:3, align = "c") |>
  style_tt(i = seq(1, 15, 2), j = 1, rowspan = 2, alignv = "t") |>
  style_tt(
    bootstrap_class = "table table-sm"
  )
```