### 2024-07-10 analysis/analysis-human-rights.qmd

---
title: "Explaining human rights violations"
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  fig.width = 6, fig.height = (6 * 0.618), out.width = "80%",
  fig.align = "center", fig.retina = 3,
  collapse = TRUE
)

options(digits = 3, width = 120, dplyr.summarise.inform = FALSE)
```

```{r}
#| label: libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(tidybayes)
library(modelsummary)
library(marginaleffects)
library(scales)
library(glue)
library(ggtext)
library(patchwork)
library(ggh4x)
library(tinytable)
library(targets)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

tar_load(c(m_hr, m_tbl_hr))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(diagnostic_functions), .GlobalEnv))
invisible(list2env(tar_read(helper_functions), .GlobalEnv))
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))

set_annotation_fonts()
```

# Model details

## Formal model specification

$$
\begin{aligned}
&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\phi_{it} =&\ \beta_0 + \beta_1\ \text{PanBack (binary)}_{it} + \beta_2\ \text{Derogation in effect}_{it} + \\
&\ \beta_3\ [\text{PanBack (binary)}_{it} \times \text{Derogation in effect}_{it}] + \\
&\ \beta_4\ \text{New cases}_{it}\ + \beta_5\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_6\ \text{New deaths}_{it}\ + \beta_7\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Week number}_{it} \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 9} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\alpha_k \sim&\ \mathcal{N}(0, 1)
\end{aligned}
$$

## Priors

```{r}
#| label: figure-prior
#| fig-cap: "Density plot of prior distribution for model parameters"
#| fig-width: 3.5
#| fig-height: 2.5
#| out-width: "50%"

ggplot() +
  stat_function(
    geom = "area",
    fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3),
    fill = clrs[2]
  ) +
  xlim(c(-20, 20)) +
  labs(x = "βs") +
  facet_wrap(vars("β: Student t(ν = 1, µ = 0, σ = 3)")) +
  theme_pandem(prior = TRUE)
```

## Simplified R code

```r
brm(
  bf(outcome ~ derogation_ineffect*panbackdichot +
      new_cases_z + cumulative_cases_z +
      new_deaths_z + cumulative_deaths_z +
      v2x_rule + year_week_num),
  family = cumulative(),
  prior = c(
    prior(student_t(1, 0, 3), class = Intercept),
    prior(student_t(1, 0, 3), class = b)),
  ...
)
```

## Model evaluation

```{r}
params_to_show <- c("b_Intercept[1]", "b_derogation_ineffect", "b_panbackdichot", "b_v2x_rule")
params_to_show_logit <- c("b_Intercept", "b_derogation_ineffect", "b_panbackdichot", "b_v2x_rule")
```

::: {.panel-tabset}

### Discriminatory policy

```{r}
plot_trace(m_hr$m_hr_discrim, params_to_show)
```

```{r}
plot_trank(m_hr$m_hr_discrim, params_to_show)
```

```{r}
plot_pp(m_hr$m_hr_discrim)
```

### Non-derogable rights

```{r}
plot_trace(m_hr$m_hr_ndrights, params_to_show_logit)
```

```{r}
plot_trank(m_hr$m_hr_ndrights, params_to_show_logit)
```

```{r}
plot_pp(m_hr$m_hr_ndrights)
```

### Abusive enforcement

```{r}
plot_trace(m_hr$m_hr_abusive, params_to_show)
```

```{r}
plot_trank(m_hr$m_hr_abusive, params_to_show)
```

```{r}
plot_pp(m_hr$m_hr_abusive)
```

### No time limits

```{r}
plot_trace(m_hr$m_hr_nolimit, params_to_show_logit)
```

```{r}
plot_trank(m_hr$m_hr_nolimit, params_to_show_logit)
```

```{r}
plot_pp(m_hr$m_hr_nolimit)
```

### Media restrictions

```{r}
plot_trace(m_hr$m_hr_media, params_to_show)
```

```{r}
plot_trank(m_hr$m_hr_media, params_to_show)
```

```{r}
plot_pp(m_hr$m_hr_media)
```

:::

# Results

## Predictions

```{r}
#| label: calc-preds-hr

# Discriminatory policy
preds_hr_discrim <- calc_preds(m_hr$m_hr_discrim)
preds_hr_discrim_details <- calc_preds_details(preds_hr_discrim)
diffs_hr_discrim <- calc_preds_diffs(preds_hr_discrim)
diffs_hr_discrim_details <- calc_preds_diffs_details(diffs_hr_discrim)

# Non-derogable rights
preds_hr_ndrights <- calc_preds(m_hr$m_hr_ndrights)
preds_hr_ndrights_plot <- bind_rows(
  preds_hr_ndrights |> mutate(.category = "Major"),
  preds_hr_ndrights |> mutate(.category = "None") |> mutate(.epred = 1 - .epred)
) |>
  mutate(.category = factor(.category, levels = c("None", "Major"), ordered = TRUE))
preds_hr_ndrights_details <- calc_preds_details(preds_hr_ndrights_plot)
diffs_hr_ndrights <- calc_preds_diffs(preds_hr_ndrights_plot)
diffs_hr_ndrights_details <- calc_preds_diffs_details(diffs_hr_ndrights)

# Abusive enforcement
preds_hr_abusive <- calc_preds(m_hr$m_hr_abusive)
preds_hr_abusive_details <- calc_preds_details(preds_hr_abusive)
diffs_hr_abusive <- calc_preds_diffs(preds_hr_abusive)
diffs_hr_abusive_details <- calc_preds_diffs_details(diffs_hr_abusive)

# Time limits
preds_hr_nolimit <- calc_preds(m_hr$m_hr_nolimit)
preds_hr_nolimit_plot <- bind_rows(
  preds_hr_nolimit |> mutate(.category = "Moderate"),
  preds_hr_nolimit |> mutate(.category = "None") |> mutate(.epred = 1 - .epred)
) |>
  mutate(.category = factor(.category, levels = c("None", "Moderate"), ordered = TRUE))
preds_hr_nolimit_details <- calc_preds_details(preds_hr_nolimit_plot)
diffs_hr_nolimit <- calc_preds_diffs(preds_hr_nolimit_plot)
diffs_hr_nolimit_details <- calc_preds_diffs_details(diffs_hr_nolimit)

# Media restrictions
preds_hr_media <- calc_preds(m_hr$m_hr_media)
preds_hr_media_details <- calc_preds_details(preds_hr_media)
diffs_hr_media <- calc_preds_diffs(preds_hr_media)
diffs_hr_media_details <- calc_preds_diffs_details(diffs_hr_media)
```

```{r}
#| label: figure-hr-preds
#| fig-cap: Predicted probabilities of violating human rights across states with low and high risks of democratic backsliding and derogation status
#| fig-width: 9.75
#| fig-height: 8.75
#| out-width: 100%
#| lightbox: true

p1 <- ggplot(preds_hr_discrim, aes(x = .draw, y = .epred)) +
  geom_area(aes(fill = .category), position = position_stack()) +
  geom_label(
    data = calc_fuzzy_labs(preds_hr_discrim_details),
    aes(x = x, y = y, label = prob_ci_nice, hjust = hjust),
    fill = scales::alpha("white", 0.4), label.size = 0,
    fontface = "bold", size = 8, size.unit = "pt"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(7, 4, 2, 1)]) +
  labs(
    x = NULL, y = "Cumulative\nprobabilities",
    fill = "Discriminatory policy", tag = "A"
  ) +
  facet_nested_wrap(
    vars(panbackdichot, derogation_ineffect),
    strip = nested_settings,
    nrow = 1
  ) +
  theme_pandem() +
  theme_fuzzy_bar

p2 <- ggplot(preds_hr_ndrights_plot, aes(x = .draw, y = .epred)) +
  geom_area(aes(fill = .category), position = position_stack()) +
  geom_label(
    data = calc_fuzzy_labs(preds_hr_ndrights_details),
    aes(x = x, y = y, label = prob_ci_nice, hjust = hjust),
    fill = scales::alpha("white", 0.4), label.size = 0,
    fontface = "bold", size = 8, size.unit = "pt"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(7, 1)]) +
  labs(
    x = NULL, y = "Cumulative\nprobabilities",
    fill = "Violation of non-derogable rights", tag = "B"
  ) +
  facet_nested_wrap(
    vars(panbackdichot, derogation_ineffect),
    strip = nested_settings,
    nrow = 1
  ) +
  theme_pandem() +
  theme_fuzzy_bar

p3 <- ggplot(preds_hr_abusive, aes(x = .draw, y = .epred)) +
  geom_area(aes(fill = .category), position = position_stack()) +
  geom_label(
    data = calc_fuzzy_labs(preds_hr_abusive_details),
    aes(x = x, y = y, label = prob_ci_nice, hjust = hjust),
    fill = scales::alpha("white", 0.4), label.size = 0,
    fontface = "bold", size = 8, size.unit = "pt"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(7, 4, 2, 1)]) +
  labs(
    x = NULL, y = "Cumulative\nprobabilities",
    fill = "Abusive enforcement", tag = "C"
  ) +
  facet_nested_wrap(
    vars(panbackdichot, derogation_ineffect),
    strip = nested_settings,
    nrow = 1
  ) +
  theme_pandem() +
  theme_fuzzy_bar

p4 <- ggplot(preds_hr_nolimit_plot, aes(x = .draw, y = .epred)) +
  geom_area(aes(fill = .category), position = position_stack()) +
  geom_label(
    data = calc_fuzzy_labs(preds_hr_nolimit_details),
    aes(x = x, y = y, label = prob_ci_nice, hjust = hjust),
    fill = scales::alpha("white", 0.4), label.size = 0,
    fontface = "bold", size = 8, size.unit = "pt"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(7, 2)]) +
  labs(
    x = NULL, y = "Cumulative\nprobabilities",
    fill = "No time limited measures", tag = "D"
  ) +
  facet_nested_wrap(
    vars(panbackdichot, derogation_ineffect),
    strip = nested_settings,
    nrow = 1
  ) +
  theme_pandem() +
  theme_fuzzy_bar

p5 <- ggplot(preds_hr_media, aes(x = .draw, y = .epred)) +
  geom_area(aes(fill = .category), position = position_stack()) +
  geom_label(
    data = calc_fuzzy_labs(preds_hr_media_details),
    aes(x = x, y = y, label = prob_ci_nice, hjust = hjust),
    fill = scales::alpha("white", 0.4), label.size = 0,
    fontface = "bold", size = 8, size.unit = "pt"
  ) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_y_continuous(labels = label_percent(), expand = c(0, 0)) +
  scale_fill_manual(values = clrs[c(7, 4, 2, 1)]) +
  labs(
    x = NULL, y = "Cumulative\nprobabilities",
    fill = "Media restrictions", tag = "E"
  ) +
  facet_nested_wrap(
    vars(panbackdichot, derogation_ineffect),
    strip = nested_settings,
    nrow = 1
  ) +
  theme_pandem() +
  theme_fuzzy_bar

layout <- "
AID
FIG
BIE
HI#
CI#
"

p1 + p2 + p3 + p4 + p5 + line_divider + line_divider + line_divider + line_divider_v +
  plot_layout(
    design = layout,
    heights = c(0.31, 0.035, 0.31, 0.035, 0.31),
    widths = c(0.94, 0.02, 0.94)
  ) +
  plot_annotation(
    caption = str_wrap(
      glue(
        "The vertical slices of the bars depict 500 posterior samples;",
        "the fuzziness represents the uncertainty in category boundaries.",
        "95% credible intervals are shown as ranges in each category",
        .sep = " "
      ),
      width = 100
    ),
    theme = theme(
      plot.caption = element_text(
        margin = margin(t = 10), size = rel(0.7),
        family = "Noto Sans", face = "plain"
      )
    )
  )
```

## Complete table of results

```{r}
#| label: table-results-full-hr
#| tbl-cap: "Complete results from models showing relationship between derogations and human rights violations (H~2~ and H~3~)"

notes <- paste(
  "Note: Estimates are median posterior log odds from logistic and ordered logistic regression models;",
  "95% credible intervals (highest density posterior interval, or HDPI) in brackets."
)

m_tbl_hr |>
  set_names(
    c("Discriminatory policy", "Non-derogable rights",
      "Abusive enforcement", "No time limits", "Media restrictions")) |>
  modelsummary(
    estimate = "{estimate}",
    statistic = "[{conf.low}, {conf.high}]",
    coef_map = coef_map,
    gof_map = gof_map,
    output = "tinytable",
    fmt = fmt_significant(2),
    notes = notes,
    width = c(0.2, rep(0.16, 5))
  ) |>
  style_tt(i = seq(1, 27, 2), j = 1, rowspan = 2, alignv = "t") |>
  style_tt(
    bootstrap_class = "table table-sm"
  )
```

## Contrasts

```{r}
#| label: figure-restriction-diffs
#| fig-cap: Contrasts in predicted probabilities of implementing COVID restrictions across states with low and high risks of democratic backsliding and derogation status
#| fig-width: 9.5
#| fig-height: 10
#| out-width: 100%
#| lightbox: true

p1 <- diffs_hr_discrim |>
  ggplot(aes(x = .epred, y = fct_rev(.category), color = .category)) +
  geom_vline(xintercept = 0, linewidth = 0.25, linetype = "21") +
  stat_pointinterval() +
  facet_nested_wrap(vars(panbackdichot, derogation_ineffect), strip = nested_settings_diffs) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(values = clrs[c(7, 4, 2, 1)], guide = "none") +
  labs(
    x = NULL, y = NULL,
    title = "Discriminatory policy", tag = "A"
  ) +
  theme_pandem() +
  theme_diffs

p2 <- diffs_hr_ndrights |>
  ggplot(aes(x = .epred, y = fct_rev(.category), color = .category)) +
  geom_vline(xintercept = 0, linewidth = 0.25, linetype = "21") +
  stat_pointinterval() +
  facet_nested_wrap(vars(panbackdichot, derogation_ineffect), strip = nested_settings_diffs) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(values = clrs[c(7, 1)], guide = "none") +
  labs(
    x = NULL, y = NULL,
    title = "Non-derogable rights", tag = "B"
  ) +
  theme_pandem() +
  theme_diffs

p3 <- diffs_hr_abusive |>
  ggplot(aes(x = .epred, y = fct_rev(.category), color = .category)) +
  geom_vline(xintercept = 0, linewidth = 0.25, linetype = "21") +
  stat_pointinterval() +
  facet_nested_wrap(vars(panbackdichot, derogation_ineffect), strip = nested_settings_diffs) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(values = clrs[c(7, 4, 2, 1)], guide = "none") +
  labs(
    x = NULL, y = NULL,
    title = "Abusive enforcement", tag = "C"
  ) +
  theme_pandem() +
  theme_diffs

p4 <- diffs_hr_nolimit |>
  ggplot(aes(x = .epred, y = fct_rev(.category), color = .category)) +
  geom_vline(xintercept = 0, linewidth = 0.25, linetype = "21") +
  stat_pointinterval() +
  facet_nested_wrap(vars(panbackdichot, derogation_ineffect), strip = nested_settings_diffs) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(values = clrs[c(7, 2)], guide = "none") +
  labs(
    x = NULL, y = NULL,
    title = "No time limited measures", tag = "D"
  ) +
  theme_pandem() +
  theme_diffs

p5 <- diffs_hr_media |>
  ggplot(aes(x = .epred, y = fct_rev(.category), color = .category)) +
  geom_vline(xintercept = 0, linewidth = 0.25, linetype = "21") +
  stat_pointinterval() +
  facet_nested_wrap(vars(panbackdichot, derogation_ineffect), strip = nested_settings_diffs) +
  scale_x_continuous(labels = label_pp) +
  scale_color_manual(values = clrs[c(7, 4, 2, 1)], guide = "none") +
  labs(
    x = NULL, y = NULL,
    title = "Media restrictions", tag = "E"
  ) +
  theme_pandem() +
  theme_diffs

(p1 + p2 + p3 + p4 + p5 + line_divider + line_divider + line_divider + line_divider_v) +
  plot_layout(
    design = layout,
    heights = c(0.31, 0.035, 0.31, 0.035, 0.31),
    widths = c(0.94, 0.02, 0.94)
  ) +
  plot_annotation(
    caption = str_wrap(
      glue(
        "Point shows posterior median;", "
        thick lines show 80% credible interval;",
        "thin black lines show 95% credible interval",
        .sep = " "
      ),
      width = 150
    ),
    theme = theme(
      plot.caption = element_text(
        margin = margin(t = 10), size = rel(0.7),
        family = "Noto Sans", face = "plain"
      )
    )
  )
```

::: {.panel-tabset}
### Discriminatory policy

```{r}
make_diffs_tbl(diffs_hr_discrim)
```

### Non-derogable rights

```{r}
make_diffs_tbl(diffs_hr_ndrights)
```

### Abusive enforcement

```{r}
make_diffs_tbl(diffs_hr_abusive)
```

### No time limits

```{r}
make_diffs_tbl(diffs_hr_nolimit)
```

### Media restrictions

```{r}
make_diffs_tbl(diffs_hr_media)
```
:::