### 2024-07-10 analysis/model-details.qmd

---
title: "Model details"
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  # echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
  fig.width = 6, fig.height = (6 * 0.618), out.width = "80%",
  fig.align = "center", fig.retina = 3,
  collapse = TRUE
)

options(digits = 3, width = 120, dplyr.summarise.inform = FALSE)
```

```{r}
#| label: libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(brms)
library(glue)
library(tinytable)
library(lubridate)
library(here)

tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

# Models
tar_load(c(m_derogations, m_restrictions, m_hr))
```

```{r calculate-model-times}
model_times <- tribble(
  ~question, ~outcome, ~family, ~model,
  "Explaining derogations", "Derogation filed", "Logistic", m_derogations$m_derogations_panback,
  "Explaining derogations", "Other treaty action", "Logistic", m_derogations$m_other_panback,
  "Explaining restrictions", "Restricted movement", "Ordered logistic", m_restrictions$m_restrict_movement,
  "Explaining restrictions", "Close public transportation", "Ordered logistic", m_restrictions$m_restrict_pubtrans,
  "Explaining restrictions", "Stay at home", "Ordered logistic", m_restrictions$m_restrict_stayhome,
  "Explaining human rights violations", "Discriminatory policy", "Ordered logistic", m_hr$m_hr_discrim,
  "Explaining human rights violations", "Non-derogable rights", "Logistic", m_hr$m_hr_ndrights,
  "Explaining human rights violations", "Abusive enforcement", "Ordered logistic", m_hr$m_hr_abusive,
  "Explaining human rights violations", "No time limits", "Logistic", m_hr$m_hr_nolimit,
  "Explaining human rights violations", "Media restrictions", "Ordered logistic", m_hr$m_hr_media
) |>
  mutate(duration = map(model, ~{
    .$fit |>
      rstan::get_elapsed_time() |>
      as_tibble() |>
      summarize(total = as.duration(max(warmup + sample)))
  })) |>
  select(-model) |>
  unnest(duration)

dur <- as.period(as.duration(sum(model_times$total)))

total_run_time <- glue(
  "{hours} hours, {minutes} minutes, and {seconds} seconds",
  hours = hour(dur), minutes = minute(dur), seconds = round(second(dur), 0)
)
```

We ran these models on a 2021 M1 MacBook Pro with 32 GB of RAM, with 4 MCMC chains spread across 8 cores, with two CPU threads per chain, using Stan through {brms} through {cmdstanr}.

In total, it took `r total_run_time` to run everything.

```{r}
total_time <- tibble(
  question = "Total", family = "—", outcome = "Total",
  total = as.duration(sum(model_times$total))
)

model_times_full <- model_times |>
  bind_rows(total_time)

group_labels_rle <- rle(model_times_full$question)
group_labels <- set_names(
  cumsum(group_labels_rle$lengths) - group_labels_rle$lengths + 1,
  group_labels_rle$values
) |> as.list()

model_times_full |>
  select(Outcome = outcome, Family = family, `Total time` = total) |>
  tt(width = c(0.4, 0.3, 0.3)) |>
  group_tt(i = group_labels) |>
  style_tt(
    i = as.numeric(group_labels) + 0:(length(group_labels) - 1),
    j = 1, bold = TRUE, background = "#e6e6e6", align = "l"
  ) |>
  style_tt(
    bootstrap_class = "table table-sm"
  )
```