### 2024-07-10 analysis/targets.qmd

---
title: "{targets} pipeline"
freeze: false
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  # echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
  fig.width = 6, fig.height = (6 * 0.618), out.width = "80%",
  fig.align = "center", fig.retina = 3,
  collapse = TRUE
)

options(digits = 3, width = 120, dplyr.summarise.inform = FALSE)
```

We use [the magical {targets} package](https://docs.ropensci.org/targets/) to run our analysis and keep track of all dependencies automatically.

To build the entire project, run `targets::tar_make()` at the R console.

Here's our complete pipeline:

```{r}
#| label: show-targets-pipeline
#| echo: false

withr::with_dir(here::here(), {
  targets::tar_glimpse()
})
```